---
title: "p8105_hw2_acm2268"
author: "Amanda Miles"
date: "10/8/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set

library(tidyverse)
library(readxl)
library(dplyr)
```

#Question 1: Mr. Trashwheel and Precipitation

##Read and clean the Mr. Trashwheel sheet

```{r}

library(readxl)
trash_data = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
      sheet = "Mr. Trash Wheel",
      skip = 1) %>%
  janitor::clean_names() %>%
  select("dumpster":"homes_powered") %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls, 0))

tail(trash_data)
```

##Read and Clean Precipitation Data for 2018 and 2019

```{r}
precip_18 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
     sheet = "2018 Precipitation",
     skip = 1) %>%
  select("Month","Total") %>%
  drop_na(Month) %>%
  mutate(year = 2018)

tail(precip_18)

precip_19 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
     sheet = "2019 Precipitation",
     skip = 1) %>%
  select("Month","Total") %>%
  drop_na(Month) %>%
    mutate(year = 2019)

tail(precip_19)

```

```{r}
precip_tidy = 
  bind_rows(precip_18, precip_19) %>%
  mutate(Month = month.name[as.numeric(Month)]) %>%
  janitor::clean_names()

print(precip_tidy)
```

#Question 2: FiveThirtyEight

## Clean data in pols-month.csv

```{r}

polls_data = read_csv("./data/pols-month.csv", col_names = TRUE) %>%
  separate(mon, sep = "-", into = c("year", "month", "day") ) %>%
  mutate(
    month = month.name[as.numeric(month)],
    president = recode(prez_dem, '0' = "gop", '1' = "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
         
print(polls_data)
```

##Clean data in snp.csv

```{r}

snp_data = read_csv("./data/snp.csv", col_names = TRUE) %>%
separate(date, sep = "/", into = c("month", "day", "year"), convert = TRUE ) %>%
  mutate(
    month = month.name[as.numeric(month)]) %>%
  mutate(
    year, year = if_else (year <= 15, year + 2000, year + 1900)) %>%
  relocate(year) %>%
  select(-day) %>%
  mutate(
     year = as.character(year))

print(snp_data)
```

##Read and clean unemployment.csv

```{r}
unempl_data = read_csv("./data/unemployment.csv", col_names = TRUE) %>%
  pivot_longer("Jan" : "Dec", names_to = "month", values_to = "unempl_percent")%>%
 mutate(
   month = recode(month, Jan = "January", Feb = "February", Mar = "March", Apr = "April", Jun = "June", Jul = "July", Aug = "August", Sept = "September", Oct = "October", Nov = "November", Dec = "December"),
   Year = as.character(Year)) %>%
  rename(year = Year)

print(unempl_data)
```

##Combine the three datasets

```{r}

polls_snp_data = left_join(polls_data, snp_data, by = c("year", "month"))

print(polls_snp_data)                   
polls_snp_un_data = left_join(polls_snp_data, unempl_data, by = c("year", "month"))

print(polls_snp_un_data)    
```

#Question 3: Popularity of Baby Names

##Load and tidy the data

```{r}
library(stringr)
names_data = read_csv("./data/Popular_Baby_Names.csv", col_names = TRUE) %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_title(childs_first_name)) %>%
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all = TRUE)
         
print(names_data)
```

##Table with the rank of the name Olivia over time

```{r}

olivia_table = filter(names_data, childs_first_name == "Olivia") %>%
  group_by(year_of_birth)

print(olivia_table, n = 24)

